Retrofitting Effect Handlers onto OCaml

### 2021-03-14 retro-concurrency.tex
done_ -> compl

\begin{lstlisting}[numbers=left]
let run main =
  let runq = Queue.create () in
	let suspend k =
	  Queue.push (fun () -> continue k ()) runq
	in
  let pending_reads = ref [] in
  let rec run_next () =
    match Queue.take_opt runq with
    | None ->
      begin match !pending_reads with
      | [] -> () (* no pending reads *)
      | todo ->
        let (done_,todo) = do_reads todo in
        List.iter (fun (str,k) ->
          Queue.push (fun () -> continue k str) runq) done_;
        pending_reads := todo;
        run_next ()
      end
    | Some f -> f ()
  in
  let rec spawn f =
    match f () with
		...
    | effect (Read ic) k ->
        pending_reads := (ic,k)::!pending_reads;
        run_next ()
  in
  ...
\end{lstlisting}

@@

When the scheduler queue is empty, the |run_next| function performs the pending
reads. We abstract away the details of the event-based I/O using the |do_reads|
function (line 13). |do_reads| takes a list of pending reads and blocks until
at least one of the reads succeeds. It returns a pair of lists |done_| and
|todo|. |done_| contains the result strings from successful reads and
corresponding continuations. These continuations are arranged to be resumed
with the read result and pushed into the scheduler queue. |todo| contains the
channels on which input is still pending and their corresponding continuations.
|pending_reads| is updated to point to the |todo| list so that they may be
attempted later. Observe that all of the changes to add asynchrony are
localised to the |run| function, and the computation that performs these
effects can remain in \emph{direct-style} (as opposed to the monadic-style in
Lwt and Async).

>>>>

\begin{lstlisting}[numbers=left]
let run main =
  let runq = Queue.create () in
	let suspend k =
	  Queue.push (fun () -> continue k ()) runq
	in
  let pending_reads = ref [] in
  let rec run_next () =
    match Queue.take_opt runq with
    | None ->
      begin match !pending_reads with
      | [] -> () (* no pending reads *)
      | todo ->
        let compl,todo = do_reads todo in
        List.iter (fun (str,k) ->
          Queue.push (fun () -> continue k str) runq) compl;
        pending_reads := todo;
        run_next ()
      end
    | Some f -> f ()
  in
  let rec spawn f =
    match f () with
		...
    | effect (Read ic) k ->
        pending_reads := (ic,k) :: !pending_reads;
        run_next ()
  in
  ...
\end{lstlisting}

@@

When the scheduler queue is empty, the |run_next| function performs the pending
reads. We abstract away the details of the event-based I/O using the |do_reads|
function (line 13). |do_reads| takes a list of pending reads and blocks until
at least one of the reads succeeds. It returns a pair of lists |compl| and
|todo|. |compl| contains the result strings from successful reads and
corresponding continuations. These continuations are arranged to be resumed
with the read result and pushed into the scheduler queue. |todo| contains the
channels on which input is still pending and their corresponding continuations.
|pending_reads| is updated to point to the |todo| list so that they may be
attempted later. Observe that all of the changes to add asynchrony are
localised to the |run| function, and the computation that performs these
effects can remain in \emph{direct-style} (as opposed to the monadic-style in
Lwt and Async).
